---
title: "Assignment 1"
author: "Briana Flynn"
date: "9/8/2020"
output: html_document
---
# Overview

For my first quant assignment, I used R studio to pull selected data from the 1-year 2018 American Community Survey ("ACS") estimates via the Public Use Microdata Sample ("PUMS") data. I was interested in understanding the prevalence of renter households in the State of MA with school-aged children without internet access, and the relationship between household internet access and limited English speaking households, household income, duration of residence in current apartment or house, and rent burden.  

## Variables Used

To explore these issues, I first installed the necessary packages and downloaded tidycensus and tidyverse to my R studio. 

```{r tidycensus, message=FALSE, warning=FALSE}
library(tidycensus)
library(tidyverse)
```

First choosing Massachusetts as my geography and renter households as my selected study group, I then chose the following variables from the ACS 1-year 2018 Survey.

  1. Household income (continuous) : HINCP
  2. Gross monthly rent payment (continuous) : GRNTP
  3. Gross rent as a percentage of monthly income (continous) : GRPIP
  4. When moved into this house or apartment (categorical) : MV
  5. Access to the internet (categorical) : ACCESS
  6. Limited English speaking household (categorical) : LNGI

In order to allow myself to filter for renter households, I also pulled the tenure variable, a categorical variable ("TEN").

## Relabeling with Mutate

The results of the "When moved into this house or apartment" variable appear in numerical categories, with digits 1 through 7 signifiying a range in the number of years the household has resided in the house or apartment which does not correspond to the digit (e.g., a result of 7 means that the household has lived there for 30 or more years). I wanted to make the results more intuitively readable, so I used mutate to convert the categories to their label in years. Some households originally had no digit assigned meaning either the household was a Group Quarters or vacant; these households show "NA" under the MV variable in my dataset. 

```{r load household-level data, results = 'hide'}
hh_data <- get_pums(variables = c("TEN", 
                                  "HINCP",
                                  "GRNTP",
                                  "MV", 
                                  "GRPIP", 
                                  "ACCESS",
                                  "HUPAC",
                                  "LNGI"),
                        state = "MA",
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(TEN_label == "Rented") %>%
  mutate(MV = case_when(
    MV == "1" ~ "0 to 1",
    MV == "2" ~ "1 to 2",
    MV == "3" ~ "2 to 4",
    MV == "5" ~ "5 to 9",
    MV == "6" ~ "10 to 19",
    MV == "7" ~ "30 or more")) %>%
  select(HINCP, GRNTP, GRPIP, MV, TEN_label, ACCESS_label, LNGI_label, HUPAC_label)

```

## Resulting Data

This pull generated results with 17,922 observations of the 8 variables of interest, meaning that the data includes 17,922 renter households in MA. The first few rows of the data frame are displaced below. 

```{r load household-level data display}
hh_data
```

The data is saved to a CSV. 

```{r save dataframes as csv}
write_csv (hh_data, "households.csv")
```
