---
title: "Assignment5"
author: "Briana Flynn"
date: "10/2/2020"
output: html_document
---

# Instructions

Estimate at least one more version of the regression you estimated in Assignment 4 by transforming one or more of the independent variables and including at least one interaction term. Which version of the model best explains the observed variation in your dependent variable?

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(jtools)
library(interactions)
library(knitr)
library(huxtable)
library(ggstance)
library(broom.mixed)
```

```{r message=FALSE, warning=FALSE}
hh_data <- read_csv("households.csv") %>%
  mutate(children = case_when(
    HUPAC_label=="No children" ~ "None",
    HUPAC_label=="With children 6 to 17 years only" ~ "6 to 17",
    HUPAC_label=="With children under 6 years and 6 to 17 years" ~ "0 to 17",
    HUPAC_label=="With children under 6 years only" ~ "0 to 6")) %>%
  mutate(internet = case_when(
    ACCESS_label=="Yes, by paying a cell phone company or Internet service provider" ~ "Yes",
    ACCESS_label=="Yes, without paying a cell phone company or Internet service provider" ~ "Yes",
    ACCESS_label=="No access to the Internet at this house, apartment, or mobile home" ~ "No")) %>%
  mutate(engl_prof = 
           LNGI_label == "At least one person in the household 14 and over speaks English only or speaks English 'very well'") %>%
  filter(HINCP > 0)
```

# Testing log-transformations and interactions to improve model fit

## Model 1: Initial Model

```{r}
options(scipen = 999)
model1 <- lm(GRPIP ~ HINCP + engl_prof + children + internet, 
            data = hh_data)

summary(model1)
```

The multiple R-squared value for the initial model is 0.2647. Two of the independent variables appear to be statistically significant: Household income, which appears to decrease rent burden by only .0001 percent for every additional dollar earned; and having no children, which appears to decrease rent burden by 2.31% when compared with households with multiple children aged 0 to 17. 


## Model 2: Log-transformed Income

```{r}
model2 <- lm(GRPIP ~ log(HINCP) + engl_prof + children + internet, 
    data = hh_data)

summary(model2)
```

When log-transforming household income, the multiple R-squared value increases to 0.3944. Now, four of my independent variables appear to be statistically significant.


## Model 3: Interactions with Log-transformed Household Income

```{r}
model3 <- lm(GRPIP ~ log(HINCP) + engl_prof + children + internet + engl_prof:log(HINCP) +
               internet:log(HINCP), 
            data = hh_data)

summary(model3)
```

When log-transforming household income and creating interaction terms between log-transformed household income and internet and English proficiency, the multiple R-squared value increases only very slightly to 0.3978. 

# Comparing the Coefficients and P-Values of Each Model

```{r}
coeff_names <- c("Household Income" = 
                   "HINCP",
                  "Household Income (log-transformed)" = 
                   "log(HINCP)", 
                 "English proficiency\n(vs. no English proficiency)" = 
                   "engl_profTRUE",
                 "Internet access" = 
                   "internetYes",
                 "Presence of children age 0 to 6" = 
                   "children0 to 6",
                 "Presence of children age 6 to 17" = 
                   "children6 to 17",
                 "No children" = 
                   "childrenNone",
                 "Interaction:\nlog-transformed income and internet access" = 
                   "log(HINCP):internetYes",
                 "Interaction:\nlog-transformed income and English proficiency" = 
                   "log(HINCP):engl_profTRUE")


export_summs(model1, model2, model3, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Model 1", "Model 2", "Preferred Model"),
             coefs = coeff_names)
```

When comparing the R-squared values of the three models, I see that the third model is my preferred model. The P-values for the second and third models are very similar.  

# Visualizing Results

## R-Squared Values by Model

```{r}
ModelFit <- tibble(model = c(1, 2, 3),
                   R_square = c(summary(model1)$adj.r.squared,
                                summary(model2)$adj.r.squared,
                                summary(model3)$adj.r.squared))

ggplot(ModelFit, aes(x = model, y = R_square)) +
  geom_line() +
  scale_x_continuous(name = "",
                   breaks = breaks <- seq(1, 4, by = 1),
                   labels = paste("Model", breaks)) +
  scale_y_continuous(name = "Adjusted R-squared value") +
  theme_bw()
```



## Household Income's Prediction of Rent Burden with Access to Internet Interaction

```{r message=FALSE, warning=FALSE}
interact_plot(model3, pred = HINCP, modx = internet, interval = TRUE) +
  scale_x_continuous(name = "Household Income")+
scale_y_continuous(name = "Rent burden",
                     limits = c(0, 100),
                     breaks = breaks <- seq(0, 100, by = 10),
                     labels = paste(prettyNum(breaks/100, big.mark = "'"),
                                    sep = "")) +
  scale_color_discrete(name = "") +
  scale_linetype_discrete(name = "") +
  scale_fill_discrete(name = "")
```

